Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
try mix-effect model (failed?)

Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis $H_0$ for departures that they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely $H_0$ is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999]. Furthermore, with a sufficiently large sample size, residual-based tests may reject $H_0$ due to a slight departure that is of little practical significance. These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.

@@

Data used in the $m - 1$ null plots need to be simulated. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggest, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

@@

As point out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. @vanderplas2021statistical summarizes three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design. For Scenario 3, @vanderplas2021statistical model the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$. Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as

@@

As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup. Thus, the power of a visual test depends on the subject-specific abilities. Previously, it was addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.

Instead, we directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as

@@

>>>>

Example residual plots given in Figure \ref{fig:residual-plot-common-departures} are examined by the corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as shown in Table \ref{tab:example-residual-plot-table}. In the example, both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null hypothesis $H_0$ for departures that they do not intend to examine. As discussed in @cook1982residuals, most residual-based tests for a particular type of departure from model assumptions are sensitive to other types of departures. It is likely $H_0$ is correctly rejected but for the wrong reason, which is known as the "Type III error". Additionally, outliers will often incorrectly trigger the rejection of $H_0$ despite when majority of the residuals are well-behaved [@cook_applied_1999]. Furthermore, with a sufficiently large sample size, residual-based tests may reject $H_0$ due to a slight departure that is of little practical significance. These can be largely avoided in diagnostic plots as experienced analysts can evaluate the acceptability of assumptions flexibly, even in the presence of outliers and slight departures.

@@

Data used in the $m - 1$ null plots need to be simulated. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

@@

As pointed out by @vanderplas2021statistical, this basic binomial model doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. @vanderplas2021statistical summarises three common scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experiment design. For Scenario 3, @vanderplas2021statistical model the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, $1/s_j$ will be added to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$. Since we are only interested in the selections of the actual data plot $i$, the marginal model can be simplified to a beta-binomial model and thus the visual p-value is given as

@@

As discussed in @majumder_validation_2013, individual's skill will affect the number of observers who identify the actual data plot from the lineup. Thus, the power of a visual test depends on the subject-specific abilities. Previously, it is addressed by modelling the probability of a subject $i$ correctly picking the actual data plot from a lineup $l$ using a mixed-effect logistic regression with the subject being treated as a random effect [@majumder_validation_2013]. However, having this probability is insufficient to determine the power of a visual test allowed for multiple selections as it doesn't provide information about the number of selections made by the subject for p-value calculation.

Instead, we omitted the individual effect and directly estimated the probability of a lineup being rejected using a logistic regression with the natural logarithm of the effect size as the only regressor formulated as

@@

```{r eval = FALSE}
MASS::glmmPQL(weighted_detect ~ log_effect_size, random = ~ log_effect_size | subject,  family = quasibinomial(), data = vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  filter(type == "polynomial") %>%
  mutate(log_effect_size = log10(effect_size)) %>%
  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
  select(log_effect_size, weighted_detect, subject)) -> mod

lme4::glmer(weighted_detect ~ log_effect_size  + (log_effect_size | subject), 
            family = binomial(),
            data = vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  filter(type == "polynomial") %>%
    mutate(log_effect_size = log10(effect_size)) %>%
  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
  select(log_effect_size, weighted_detect, subject)) -> mod

vi_survey %>%
  filter(!attention_check) %>%
  filter(!null_lineup) %>%
  filter(x_dist == "uniform") %>%
  filter(type == "polynomial") %>%
  mutate(log_effect_size = log10(effect_size)) %>%
  mutate(subject = factor(as.integer(factor(paste(exp, set))))) %>%
  mutate(pred_wd = predict(mod, type = "response")) %>%
  group_by(unique_lineup_id) %>%
  mutate(pred_ci = pred_wd * n()) %>%
  mutate(n_eval = n()) %>%
  ungroup() %>%
  rowwise() %>%
  mutate(pred_pval = (function(c_i, n_eval, alpha) {
  target_dist <- exact_dist(n_eval, 20, dist = "dirichlet", alpha = alpha)
  floor_c_i <- floor(c_i)
  ceil_c_i <- ceiling(c_i)
  return(unname(sum(target_dist[(ceil_c_i:n_eval) + 1]) + (ceil_c_i - c_i) * target_dist[floor_c_i + 1]))
})(pred_ci, n_eval, alpha)) %>%
  ungroup() %>%
  mutate(pred_reject = pred_pval < 0.05) %>%
  ggplot() +
  geom_smooth(aes(log_effect_size, as.numeric(pred_reject)), method = "glm", method.args = list(family = binomial()), se = FALSE)

```

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
try mix-effect model (failed?)

Example residual plots given in Figure
\ref{fig:residual-plot-common-departures} are examined by the
corresponding RESET test, Breusch-Pagan test and Shapiro--Wilk test as
shown in Table \ref{tab:example-residual-plot-table}. In the example,
both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
hypothesis \(H_0\) for departures that they do not intend to examine. As
discussed in \citet{cook1982residuals}, most residual-based tests for a
particular type of departure from model assumptions are sensitive to
other types of departures. It is likely \(H_0\) is correctly rejected
but for the wrong reason, which is known as the ``Type III error''.
Additionally, outliers will often incorrectly trigger the rejection of
\(H_0\) despite when majority of the residuals are well-behaved
\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
size, residual-based tests may reject \(H_0\) due to a slight departure
that is of little practical significance. These can be largely avoided
in diagnostic plots as experienced analysts can evaluate the
acceptability of assumptions flexibly, even in the presence of outliers
and slight departures.

@@

Data used in the \(m - 1\) null plots need to be simulated. In the
context of regression diagnostics, sampling data from \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja_statistical_2009} suggest, \(H_0\) is usually composited by
a collection of distributions controlled by nuisance parameters. Since
regression models can have various forms, there is no general solution
to this problem, but it sometimes can be reduced to so called
``reference distribution'' by applying one of the three methods: (i)
sampling from a conditional distribution given a minimal sufficient
statistic under \(H_0\), (ii) parametric bootstrap sampling with
nuisance parameters estimated under \(H_0\), and (iii) Bayesian
posterior predictive sampling. The conditional distribution given a
minimal sufficient statistic is the best justified reference
distribution among the three \citep{buja_statistical_2009}. Essentially,
null residuals can be simulated by regressing \(N\) i.i.d standard
normal random draws on the regressors, then rescaling it by the ratio of
residual sum of square in two regressions.

@@

As point out by \citet{vanderplas2021statistical}, this basic binomial
model doesn't take into account the possible dependencies in the visual
test due to repeated evaluations of the same lineup. And it is
inapplicable to visual test where subjects are asked to select one or
more ``most different'' plots from the lineup.
\citet{vanderplas2021statistical} summarizes three common scenarios in
visual inference: (1) \(K\) different lineups are shown to \(K\)
subjects, (2) \(K\) lineups with different null plots but the same
actual data plot are shown to \(K\) subjects, and (3) the same lineup is
shown to \(K\) subjects. Out of these three scenarios, Scenario 3 is the
most common in previous studies as it puts the least constraints on the
experiment design. For Scenario 3, \citet{vanderplas2021statistical}
model the probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
instead of one, where \(s_j\) is the number of plots subject \(j\)
selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Since we are
only interested in the selections of the actual data plot \(i\), the
marginal model can be simplified to a beta-binomial model and thus the
visual p-value is given as

@@

As discussed in \citet{majumder_validation_2013}, individual's skill
will affect the number of observers who identify the actual data plot
from the lineup. Thus, the power of a visual test depends on the
subject-specific abilities. Previously, it was addressed by modelling
the probability of a subject \(i\) correctly picking the actual data
plot from a lineup \(l\) using a mixed-effect logistic regression with
the subject being treated as a random effect
\citep{majumder_validation_2013}. However, having this probability is
insufficient to determine the power of a visual test allowed for
multiple selections as it doesn't provide information about the number
of selections made by the subject for p-value calculation.

Instead, we directly estimated the probability of a lineup being
rejected using a logistic regression with the natural logarithm of the
effect size as the only regressor formulated as

>>>>

Example residual plots given in Figure
\ref{fig:residual-plot-common-departures} are examined by the
corresponding RESET test, Breusch-Pagan test and Shapiro-Wilk test as
shown in Table \ref{tab:example-residual-plot-table}. In the example,
both the Breusch-Pagan test and the Shapiro--Wilk test rejects the null
hypothesis \(H_0\) for departures that they do not intend to examine. As
discussed in \citet{cook1982residuals}, most residual-based tests for a
particular type of departure from model assumptions are sensitive to
other types of departures. It is likely \(H_0\) is correctly rejected
but for the wrong reason, which is known as the ``Type III error''.
Additionally, outliers will often incorrectly trigger the rejection of
\(H_0\) despite when majority of the residuals are well-behaved
\citep{cook_applied_1999}. Furthermore, with a sufficiently large sample
size, residual-based tests may reject \(H_0\) due to a slight departure
that is of little practical significance. These can be largely avoided
in diagnostic plots as experienced analysts can evaluate the
acceptability of assumptions flexibly, even in the presence of outliers
and slight departures.

@@

Data used in the \(m - 1\) null plots need to be simulated. In the
context of regression diagnostics, sampling data from \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja_statistical_2009} suggested, \(H_0\) is usually composited
by a collection of distributions controlled by nuisance parameters.
Since regression models can have various forms, there is no general
solution to this problem, but it sometimes can be reduced to so called
``reference distribution'' by applying one of the three methods: (i)
sampling from a conditional distribution given a minimal sufficient
statistic under \(H_0\), (ii) parametric bootstrap sampling with
nuisance parameters estimated under \(H_0\), and (iii) Bayesian
posterior predictive sampling. The conditional distribution given a
minimal sufficient statistic is the best justified reference
distribution among the three \citep{buja_statistical_2009}. Essentially,
null residuals can be simulated by regressing \(N\) i.i.d standard
normal random draws on the regressors, then rescaling it by the ratio of
residual sum of square in two regressions.

@@

As pointed out by \citet{vanderplas2021statistical}, this basic binomial
model doesn't take into account the possible dependencies in the visual
test due to repeated evaluations of the same lineup. And it is
inapplicable to visual test where subjects are asked to select one or
more ``most different'' plots from the lineup.
\citet{vanderplas2021statistical} summarises three common scenarios in
visual inference: (1) \(K\) different lineups are shown to \(K\)
subjects, (2) \(K\) lineups with different null plots but the same
actual data plot are shown to \(K\) subjects, and (3) the same lineup is
shown to \(K\) subjects. Out of these three scenarios, Scenario 3 is the
most common in previous studies as it puts the least constraints on the
experiment design. For Scenario 3, \citet{vanderplas2021statistical}
model the probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
instead of one, where \(s_j\) is the number of plots subject \(j\)
selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Since we are
only interested in the selections of the actual data plot \(i\), the
marginal model can be simplified to a beta-binomial model and thus the
visual p-value is given as

@@

As discussed in \citet{majumder_validation_2013}, individual's skill
will affect the number of observers who identify the actual data plot
from the lineup. Thus, the power of a visual test depends on the
subject-specific abilities. Previously, it is addressed by modelling the
probability of a subject \(i\) correctly picking the actual data plot
from a lineup \(l\) using a mixed-effect logistic regression with the
subject being treated as a random effect
\citep{majumder_validation_2013}. However, having this probability is
insufficient to determine the power of a visual test allowed for
multiple selections as it doesn't provide information about the number
of selections made by the subject for p-value calculation.

Instead, we omitted the individual effect and directly estimated the
probability of a lineup being rejected using a logistic regression with
the natural logarithm of the effect size as the only regressor
formulated as
